---
title: "Squirrel"
output: html_notebook
---

```{r}
# Load packages
library(raster)
library(leaflet)
library(plotly)
library(igraph)
library(Matrix)

# Load homemade functions
source('build_matrix_graph.R')
```


# Load raster

The raster is freely available from the web

```{r, echo=TRUE}
# Prepare globCover dataset
# download.file('http://due.esrin.esa.int/files/Globcover2009_V2.3_Global_.zip',destfile = "Globcover2009_V2.3_Global_.zip")
# unzip('Globcover2009_V2.3_Global_.zip',exdir = 'GlobCover')
```

We load the Spanish borders and plot the raster. Also, we map the Globcover labels to a continous number, as indicated by the legend found [here](http://due.esrin.esa.int/page_globcover.php). We set water bodies to _NA_, to indicate that the squirrel cannot go over them.
```{r}
# Crop the raster to view Spain only
esp_border <- getData("GADM",country="Spain",level=0)
glob = raster("GLOBCOVER_L4_200901_200912_V2.3.tif")
glob_esp_orig = crop(glob,esp_border)

## ================
# Transform labels to roughness values
globCover = c(11,14,20,30,40,50,60,70,90,100,
              110,120,130,140,150,160,170,180,190,200,210,220,230)
roughness = c(0.10,  0.10,  0.30,  0.30,  1.50,  1.50,  1.50,  1.50,  1.50,  1.50,
              1.50,  0.50,  0.10,  0.03,  0.05,  0.50,  0.60,  0.20,  1.00,  NA,  NA,  NA,  NA)
glob_esp_plot = reclassify(glob_esp_orig,cbind(globCover,roughness))
# Let's tranform NA values to "-100" roughness so that we squirrel cannot swim. This one is only used for calculations.
glob_esp = reclassify(glob_esp_plot,c(NA,-100,-100))

par(mfrow=c(1,2))
plot(glob_esp_plot)
plot(glob_esp)
```

```{r}
## ================
# pal2 = colorNumeric("Greens", domain = NULL)
# # Cretae map
# ll_map =  leaflet() %>%
#   # Base maps
#   addProviderTiles('Esri.WorldTopoMap',group="Esri Topo") %>%
#   addProviderTiles("Esri.WorldImagery", group = "Esri Image") %>%
#   # Roughness info
#   addRasterImage(glob_esp, opacity = 0.5,colors=pal2, group="Roughness map") %>%
# 
#   addPolygons(data = esp_border,
#               fill = F, weight = 2, color = "black", group = "Borders") %>%
#   addLegend(position="bottomleft", pal = pal2, values = values(glob_esp),
#             title = "Roughness",
#             opacity = 1) %>%
#   # Control groups
#   addLayersControl(
#     baseGroups = c("Esri Image","Esri Topo"),
#     overlayGroups = c("Roughness map", "Borders"),
#     options = layersControlOptions(collapsed = F)
#   )
# ll_map
```


# The path

For testing purposes we start by chopping a small section
```{r}
# Chop a small section
lat = 43.05
lon = -4.05
n=0.15 # width of the box
cropbox2 <-c(lon-n,lon+n,lat-n,lat+n)
small_img = crop(glob_esp,cropbox2)
small_img_plot = crop(glob_esp_plot,cropbox2)
# Plot
par(mfrow=c(1,2))
plot(small_img_plot)
plot(small_img)
# translate raster to matrix
mat = raster::as.matrix(small_img)
dim(mat)
```



```{r}
# create 0-1 edge matrix
N = dim(mat)[1] * dim(mat)[2] # Number of vertices
Ni= dim(mat)[1] # number of rows
Nj= dim(mat)[2] # number of columns

# Call our homemade function
mat_path = build_matrix_graph(Ni,Nj)


# Add weights to each edge
for(i in 1:N){
  if( is.na(values(small_img)[i])){
    mat_path[i,mat_path[i,] != 0] = 0
  }else{
    mat_path[i,mat_path[i,] != 0] = 0.001 + 1.5 - values(small_img)[i]
  }
}

# Build igraph
g <- graph.adjacency(mat_path, mode="directed",weighted = T)
  
# Calculate shortest path between two poitns
p1 = c(1,2)
p2 = c(Ni,Nj-1)
news.path <- shortest_paths(g,from = index2num(p1[1],p1[2],Nj), 
                            to  = index2num(p2[1],p2[2],Nj),
                            output = "both") # both path nodes and edges

# Find those poitns in lat,lon
path_squirrel = news.path$vpath[[1]] %>% as.numeric
path_points = rasterToPoints(small_img)[path_squirrel,1:2]
```

### Visualize solution

```{r}
# Make a leaflet visualization
ll_map =  leaflet() %>%
  # Base maps
  addProviderTiles('Esri.WorldGrayCanvas',group="Esri Grey") %>%
  addProviderTiles('Esri.WorldTopoMap',group="Esri Topo") %>%
  addProviderTiles("Esri.WorldImagery", group = "Esri Image") %>%
  # Roughness info
  addRasterImage(small_img_plot, opacity = 0.5,colors=pal2, group="Roughness map") %>%
  addLegend(position="bottomleft", pal = pal2, values = values(small_img_plot),
            title = "Roughness",
            opacity = 1) %>%
  # path info
  addPolylines(lng=path_points[,1],lat=path_points[,2]) %>% 
  # Control groups
  addLayersControl(
    baseGroups = c("Esri Grey","Esri Image","Esri Topo"),
    overlayGroups = c("Roughness map", "Borders"),
    options = layersControlOptions(collapsed = F)
  )
ll_map
# library(htmlwidgets)
# saveWidget(ll_map,'example.html')
```

### Some statistics on the path

Average roughness, % numer of steps outside trees..


